---
title: "Operators in R"
output:
  html_document:
    df_print: paged
---


- Arithmetic Operators

- Relational Operators

- Logical Operators

- Assignment Operators


# Arithmetic/Numeric operators

used with numeric data types to perform mathematical operations

- +: addition

- -: subtraction

- *: multiplication

- /: division

- Modulo Operator (%%):

- ^: raising to power

## Addition 
```{r}
# set values of x, y and z for numeric operators exploration
x <- 5   # integer value
y <- 2.0 # float value
z <- -13 # negative integer

#(run this cell to initialize values for the next few cells!)
```

```{r}
# sum of x and y
print(x + y)
```
```{r}
# Adding two vectors
p<- c(5,6)
q<- c(7,8)
print(p+q)
```
## Subtraction Operator (-): 
```{r}
# difference of x and y
print(x - y)
```

```{r}
# Subtraction of two vectors
p<- c(5,6)
q<- c(7,8)
print(p-q)
print(q-p)


```
## Multiplication

```{r}
# product of x and y
print(x * y)
```

```{r}
# vector multiplication
print(p*q)
```
## Division
```{r}
# quotient of x and y
print(x / y)
```

## Power Operator (^)
```{r}
# x raised to power y
print(x ^ y)
```
## Modulo Operator (%%):
```{r}
print(x %% y)
```


## Excercise 

Create a variable with the value of pi as 3.14, create a new variable as radius and assign it a float value, this will be the radius of a circle and then output the area of the circle using the formula pi*radius^2


# Comparison Operators

used to compare two values

- (<): less than

- (>): greater than

- (<=): less than or equal to

- (>=): greater than or equal to

- (==): equality

- (!=): inequality


```{r}
# Greater Than 
i = 3 
j = 7

print(i > j) 
```

```{r}
# Greater Than or Equal
i = 3 
j = 7

print(i >= j) 
```

```{r}
# Less Than 
i = 3 
j = 7

print(i < j) 
```

```{r}
# Less Than or Equal 
i = 3 
j = 7

print(i <= j) 
```

```{r}
# Equality Test #1
i = 3
j = 7

print(i == j)
```

```{r}
# Equality Test #2
i = 3
j = 3
print(i == j)
```
```{r}
# Inequality Test #1
i = 3
j = 7

print(i != j)
```


```{r}
# Inequality Test #2
i = 3
j = 3

print(i != j)
```

## Excercise :

In the last excercise you created a program to get the area of a circle , create a print function that checks if the area that your program created is bigger than 50 or not

# Logic Operators

used to combine conditional statements
- And = &
- Or = |
- Not = !

**Example :-**
& :

```{r}
v <- c(3,1,TRUE,2+3i)
t <- c(4,1,FALSE,2+3i)
print(v&t)
```


**|** :

```{r}
v <- c(3,0,TRUE,2+2i)
t <- c(4,0,FALSE,2+3i)
print(v|t)
 
```

 **!** :
 
```{r}
v <- c(3,0,TRUE,2+2i)
print(!v)
```

it produces the following output :-

 
## Assignment Operators

These operators are used to assign values to vectors.

**Left Assignment : <- or = or <<- :**

**Right Assignment : -> or = or ->> :**

Example :

**Left assignment <-, <<-, =**

```{r}
v1 <- c(3,1,TRUE,2+3i)
v2 <<- c(3,1,TRUE,2+3i)
v3 = c(3,1,TRUE,2+3i)
print(v1)
print(v2)
print(v3)
```

 
**Right assignment -> , ->> , =**


```{r}
c(3,1,TRUE,2+3i) -> v1
c(3,1,TRUE,2+3i) ->> v2 
print(v1)
print(v2)
```



